---
title: "CodeBook"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data

The data for this project comes from the "Human Activity Recognition Using Smartphones Dataset - Version 1.0", available through the linked [UCI HAR Dataset](https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip). Information about how this data was collected can be found at the [Human Activity Recognition Using Smartphones](http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones) website. In summary, the data comes from an experiment of 30 volunteers between the ages of 19-48. The participants performed 6 different activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) while wearing a Samsun Galaxy S II and then accelerometer and gyroscope data was collected. 

Specifically, for each participant, the following information was collected as raw, time domain signals:

* Triaxial acceleration from the accelerometer (total acceleration) and the estimated body acceleration.
* Triaxial Angular velocity from the gyroscope. 

The time domain signals were captured at a constant rate of 50 Hz and then denoised by filtration through both a median and a 3rd order low pass Butterworth filter, with corner frequency of 20Hz. The acceleration signal, after being sepeartaed into body and gravity signals was also filtered using a low pass Butterworth filter with corner frequency 0.3 Hz. 


The citation for this dataset is: 

[1] Davide Anguita, Alessandro Ghio, Luca Oneto, Xavier Parra and Jorge L. Reyes-Ortiz. Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine. International Workshop of Ambient Assisted Living (IWAAL 2012). Vitoria-Gasteiz, Spain. Dec 2012


## Variables
Measurements were takenf or each of the 30 participants and each of the 6 activities. The entries under each variable name is the mean of the total measurements of that type that were taken for the specific subject and activity noted in columns 1 and 2. 

All variable names (columns 3 - 87) are made up of five components, which explain which signal comprised each measurement: 

1. *Domain of the Signal* 
    a. *Time*  = the signal is in the time doman 
    b. *Freq* = a Fast Fourier Transform was applied to put the signal in the frequency domain
2. *Component of the Signal*
    a. *Body* = the measureement is the body motion component of the signal
    b. *Gravity* = the measurement is the gravitational component of the signal
3. *Type of Signal*
    a. *Acc* = acceleration 
    b. *AccJerk* = Jerk (obtained by differentiang the body linear acceleration in time)
    d. *Gyro* = reading from the gyroscope (angular velocity)
    e. *GyroJerk* = Jerk (obtained by differenting the angular velocity in time)
    f. *AccMag* = the magnitude of the acceleration signal
    g. *GyroMag* = the magnitude ofthe gyroscope signal
4. *Type of Measurment*
    a. *Mean* = the mean of the signal
    b. *Std* = the standard deviation of the signal
3. *Direction of the Signal*
    a. *X* = the axial signal in the X direction
    b. *Y* = the axial signal in the Y direction
    c. *Z* = the axial signal in the Z direction


## Data Cleaning

The cleaning of the data to create tidydata.csv was comprised of four major steps. 

**Merging** 
In the original dataset, 70% of the subjects generated the training data and the remaining 30% generated the test data. The goal in this section was to merge these two datasets into one dataset.

To each of the training set (X_train.txt) and the testing set (X_test.txt), two additional columns were added to the leftmost part of the dataframe, to identify the subject (taken from subject_train.txt and subject_test.txt, respectively) and the activity performed (taken from y_train.txt. and y_test.txt, respectively) for each row of measurements. 

The two datasets were then merged using a row bind. 

**Extracting Desired Columns**
Only variables which measured the mean and standard deviation were kept in the tidied dataset. All other columns were removed. This was accomplished with the following commands: 

```{r, eval = FALSE}
colInds <- grep("mean|std",features$V2, ignore.case = TRUE)
subset_data <- merged_data[,c(1:2,colInds+2)] 
```

**Relabeling**
In the original dataset, the six acitvities that were measured were identified only by a number (e.g. 1 = walking). First, the activity numerical labels were replaced with descriptive lables as given in activity_labels.txt. 

```{r, eval = FALSE}
activity_labels <- read.table('../UCI HAR Dataset/activity_labels.txt')
subset_data$activity <- activity_labels[match(subset_data$activity, activity_labels$V1),]$V2
```

Then, the feature (variable) names were cleaned to be made more descriptive, by removing puncuation and expanding the prefixes "t" and "f" to be "Time" and "Freq", respectively. 

**Aggregating**
Finally, all rows with the same Subject and Activity label (columns 1 and 2, respectively) were aggregated together by taking the mean of the corresponding measurements. 
```{r, eval = FALSE}
tidy_data = aggregate(subset_data, by= list(subset_data$subject,subset_data$activity),FUN = mean)
```

The rows were then sorted in chronological order of subject and exported to a .csv file. 
